---
title: "The BILJOU forest water balance model in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The BILJOU forest water balance model in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

`biljouR` is an independent R re-implementation of the **BILJOU** lumped daily
forest water balance model (Granier, Bréda, Biron & Villette, 1999, *Ecological
Modelling* 116:269-283), as documented by INRAE UMR Silva
(<https://appgeodb.nancy.inrae.fr/biljou/>). It is **not** produced or endorsed
by INRAE; use the official tool for authoritative simulations.

## The model in one equation

The model integrates, at a **daily** time step, the soil water balance

$$\Delta W = P - In - T - Eu - D$$

where `P` is rainfall, `In` rainfall interception, `T` overstorey
transpiration, `Eu` understorey + soil evaporation and `D` drainage. Drainage
is obtained by difference / cascade through the soil layers.

### Elementary fluxes

* **LAI phenology.** Evergreen stands keep a constant LAI; deciduous stands ramp
  from 0 to `lai_max` over 30 days after budburst and back to 0 over the 30 days
  before leaf fall.
* **Transpiration.** `T = r(LAI) · PET`, with `r = 0.125·LAI` for `LAI ≤ 6` and
  `r = 0.75` above. `T` is reduced by 20 % of intercepted water (wet canopy) and
  scaled by `REW/REWc` under water stress (`REWc = 0.4`).
* **Interception.** `Th = exp(a + b·R/R0 + c·P + d·P²)`, `In = P − Th`
  (Aussenac 1968), with stand-type coefficients; `R/R0 = 100·exp(−k·LAI)`.
* **Understorey/soil evaporation.** `Eu = PET·exp(−k·LAI)`, limited by the
  upper-layer water.
* **Soil & drainage.** Each layer refills through its microporosity while a
  fraction bypasses quickly through macroporosity; water above field capacity
  drains to the next layer. Root uptake is weighted by fine-root fraction and
  current relative extractable water.

## A worked example

```{r}
library(biljouR)
data(meteo_hesse)

soil <- biljou_soil(
  ewm   = c(70, 70, 40),                 # mm extractable water per layer
  roots = c(0.6, 0.3, 0.1),              # fine-root fractions
  macro = c(0.3, 0.2, 0.1),              # optional macroporosity
  micro = c(0.7, 0.8, 0.9)               # optional microporosity
)
soil

run <- biljou_run(meteo_hesse, soil, lai_max = 6,
                  forest_type = "broadleaved",
                  budburst = 110, leaf_fall = 300)
run
```

### Drought indices and balance

```{r}
biljou_indices(run)
biljou_annual_balance(run)
```

### Visualising the season

```{r}
plot(run)
```

## Computing Penman PET

BILJOU is calibrated on **Penman** PET. If you do not already have it:

```{r}
penman_pet(tmean = 18, rg = 22, wind = 2, rh = 65,
           doy = 180, latitude = 48.6, altitude = 250)
```

## Caveats

Some constants are only described qualitatively in the public BILJOU
documentation. The choices made here (macro/micro split, root-uptake weighting,
understorey coefficient, interception clamping) are transparent and documented
in the source, but should be **validated against the official tool** before
production use.
